### 2021-12-11 thesis/tex/oscillations.tex
update

\subsection{Three Neutrino Picture}
Reminding ourselves that the vacuum Hamiltonian $H_0$ has eigenstates in the mass basis, we write the following expression where 
use the relations Eq.~\ref{eq:osc_1} and Eq.~\ref{eq:osc_k} to switch between the flavor and mass basis with the PMNS elements:
\begin{align}
    \bra{\nu_\b} H_0 &= \sum_k U_{\b k} \bra{\nu_k}H_0 \nonumber \\
                     &= \sum_k U_{\b k} E_k \bra{\nu_k} \nonumber \\
                     &= \sum_\eta \sum_k U_{\b k} E_k U^*_{\eta k} \bra{\nu_\eta}\,.
\end{align}
Thus,
\begin{align}
    \bra{\nu_\b}H_0\ket{\nu_\a (t)} &= \sum_\eta \sum_k U_{\b k} E_k U^*_{\eta k} \bra{\nu_\eta}\ket{\nu_\a (t)} \nonumber \\
                                    &= \sum_\eta \sum_k U_{\b k} E_k U^*_{\eta k} \psi_{\a\eta}(t)\,.
\end{align}
Using the ultrarelativistic approximation from Eq.~\ref{eq:ultra_rel}:
\begin{align}
    \sum_\eta \sum_k U_{\b k} E_k U^*_{\eta k} \psi_{\a\eta}(t) &= \sum_\eta \sum_k U_{\b k} \left(p + \frac{m^2_k}{2E}\right) U^*_{\eta k} \psi_{\a\eta}(x) \nonumber \\
    &= \sum_\eta \sum_k U_{\b k} \left(p + \frac{m^2_k}{2E}\right) U^*_{\eta k} \psi_{\a\eta}(x)\,.
\end{align}
Use the fact that $\sum_k m_k^2 =  \sum m_1^2 +_k m^2_k - m^2_1 =\sum_k m_1^2 +  \dm_{k1}$ to pull out common terms out of the summation:
\begin{align}\label{eq:t1}
    \sum_\eta \sum_k U_{\b k} \left(p + \frac{m^2_k}{2E}\right) U^*_{\eta k} \psi_{\a\eta}(x) &= \sum_\eta \sum_k U_{\b k} \left(p + \frac{m^2_1}{2E} + \frac{\dm_{k1}}{2E}\right) U^*_{\eta k} \psi_{\a\eta}(x) \nonumber \\
    &= \sum_\eta\sum_k \left(p + \frac{m^2_1}{2E}\right) U_{\b k} U^*_{\eta k} \psi_{\a\eta}(x) \nonumber \\
    &+ \sum_\eta\sum_k U_{\b k} \frac{\dm_{k1}}{2E} U^*_{\eta k} \psi_{\a\eta}(x)\,.
\end{align}
Unitarity gives $ \sum_k U_{\b k} U^*_{\eta k} = \delta_{\eta\b}$, and the first term in the last step of Eq.~\ref{eq:t1} becomes
\begin{align}
    \sum_\eta \left(p + \frac{m^2_1}{2E}\right) \delta_{\beta \eta} \psi_{\a\eta}(x)
    =& \left(p + \frac{m^2_1}{2E}\right) \psi_{\a\b}(x)\,.
\end{align}

@@

\subsection{Two Neutrino Picture Solution}
By working in the limit where one of the neutrino masses can be neglected, 
we can reduce the complexity of the neutrino oscillations and obtain tractable analytical 
expressions. 
We now only consider the mixing between two neutrino flavors, say between $\ne, \nm$, and $\nu_1, \nu_2$. In other words, we let $\dm[31] = 0$. For simplicity, let $\delta_{CP} = 0$.

@@

\begin{align}
    P_{e\mu}^M = \sin^2(2\theta_{12}^M)\sin^2\left(\frac{\dm[21]L}{4E}\right)\,.
\end{align}

@@

\subsection{Earth propagation}
We now need to know how the electrons are distributed within the Earth. The Preliminary Earth Reference Model~\cite{PREM} gives us spherically
symmetric piecewise polynomials for the Earth density in \si{\gram\cm^{-3}} shown in the left panel of Fig.~\ref{fig:potential}.
We note a steep discontinuity at \SI{3480}{\km} where the density is nearly halved. This is the core-mantle boundary and will be visible in our 
oscillations.

>>>>

\subsection{The Three Neutrino Picture}
Reminding ourselves that the vacuum Hamiltonian $H_0$ has eigenstates in the mass basis, we write the following expression where 
use the relations Eq.~\ref{eq:osc_1} and Eq.~\ref{eq:osc_k} to switch between the flavor and mass basis with the PMNS elements:
\begin{align}
    \bra{\nu_\b} H_0 &= \sum_k U_{\b k} \bra{\nu_k}H_0 \nonumber \\
                     &= \sum_k U_{\b k} E_k \bra{\nu_k} \nonumber \\
                     &= \sum_\eta \sum_k U_{\b k} E_k U^*_{\eta k} \bra{\nu_\eta}\,.
\end{align}
Thus,
\begin{align}
    \bra{\nu_\b}H_0\ket{\nu_\a (t)} &= \sum_\eta \sum_k U_{\b k} E_k U^*_{\eta k} \bra{\nu_\eta}\ket{\nu_\a (t)} \nonumber \\
                                    &= \sum_\eta \sum_k U_{\b k} E_k U^*_{\eta k} \psi_{\a\eta}(t)\,.
\end{align}
Using the ultrarelativistic approximation from Eq.~\ref{eq:ultra_rel}:
\begin{align}
    \sum_\eta \sum_k U_{\b k} E_k U^*_{\eta k} \psi_{\a\eta}(t) &= \sum_\eta \sum_k U_{\b k} \left(p + \frac{m^2_k}{2E}\right) U^*_{\eta k} \psi_{\a\eta}(x) \nonumber \\
    &= \sum_\eta \sum_k U_{\b k} \left(p + \frac{m^2_k}{2E}\right) U^*_{\eta k} \psi_{\a\eta}(x)\,.
\end{align}
Use the fact that $\sum_k m_k^2 =  \sum m_1^2 +_k m^2_k - m^2_1 =\sum_k m_1^2 +  \dm_{k1}$ to pull out common terms out of the summation:
\begin{align}\label{eq:t1}
    \sum_\eta \sum_k U_{\b k} \left(p + \frac{m^2_k}{2E}\right) U^*_{\eta k} \psi_{\a\eta}(x) &= \sum_\eta \sum_k U_{\b k} \left(p + \frac{m^2_1}{2E} + \frac{\dm_{k1}}{2E}\right) U^*_{\eta k} \psi_{\a\eta}(x) \nonumber \\
    &= \sum_\eta\sum_k \left(p + \frac{m^2_1}{2E}\right) U_{\b k} U^*_{\eta k} \psi_{\a\eta}(x) \nonumber \\
    &+ \sum_\eta\sum_k U_{\b k} \frac{\dm_{k1}}{2E} U^*_{\eta k} \psi_{\a\eta}(x)\,.
\end{align}
Unitarity gives $ \sum_k U_{\b k} U^*_{\eta k} = \delta_{\eta\b}$, and the first term in the last step of Eq.~\ref{eq:t1} becomes
\begin{align}
    \sum_\eta \left(p + \frac{m^2_1}{2E}\right) \delta_{\beta \eta} \psi_{\a\eta}(x)
    =& \left(p + \frac{m^2_1}{2E}\right) \psi_{\a\b}(x)\,.
\end{align}

@@

\subsection{Earth Propagation}
We now need to know how the electrons are distributed within the Earth. The Preliminary Earth Reference Model~\cite{PREM} gives us spherically
symmetric piecewise polynomials for the Earth density in \si{\gram\cm^{-3}} shown in the left panel of Fig.~\ref{fig:potential}.
We note a steep discontinuity at \SI{3480}{\km} where the density is nearly halved. This is the core-mantle boundary and will be visible in our 
oscillations.

\begin{figure}
    \centering
    \includegraphics[width=0.9\textwidth]{figures/vac_vs_matter.png}
    \caption{$\nm$ oscillation probabilities after traversing the diameter of the Earth in black, 
    compared to the vacuum oscillation with the same 
    baseline in red. 
    Oscillation parameters are taken from NuFit~\cite{nufit} and given in Eq.~\ref{eq:PINGUparams}.}\label{fig:vac_vs_matter}
\end{figure}

In Fig.~\ref{fig:vac_vs_matter}, we have plotted the oscillation probabilities for an incoming $\nm$ with \si{\GeV} energy,
traveling 12000 km. The red line is the vacuum oscillation, while the black line is the matter oscillation, traversing through the 
matter density profile of the Earth. We see that the matter effects are more for some energies, and less for others. At energies 
above $\SI{10}{\GeV}$, the matter effects are small for $\Pmm$ and $\Pmt$, while they heavily deviate for single-digit \si{\GeV} energies.
$\Pme$ behaves in a different way, where the matter effects affect the oscillation probability regardless of energy. Here, we start to
see the convergence as late as $\SI{100}{\GeV}$. Thus, matter effects are most important for energies below $\SI{100}{\GeV}$, for all channels.
Above these energies, the density profile of the Earth is less relevant, and the Earth is more or less transparent for $\nm$.

\subsection{The Two Neutrino Picture}
By working in the limit where one of the neutrino masses can be neglected, 
we can reduce the complexity of the neutrino oscillations and obtain tractable analytical 
expressions. 
We now only consider the mixing between two neutrino flavors, say between $\ne, \nm$, and $\nu_1, \nu_2$. In other words, we let $\dm[31] = 0$. For simplicity, let $\delta_{CP} = 0$.

@@

\begin{align}\label{eq:2gen}
    P_{e\mu}^M = \sin^2(2\theta_{12}^M)\sin^2\left(\frac{\dm[21]L}{4E}\right)\,.
\end{align}

@@

\begin{figure}
    \centering
    \includegraphics[width=1\textwidth]{figures/2_vs_3.png}
    \caption{$\nm$ oscillation probabilities after traversing the diameter of the Earth in black, 
    compared to the two-neutrino picture oscillations with the same 
    baseline and matter potential in blue. The figure showcases the convergence of the $2\nu$ and $3\nu$ pictures at energies above $\SI{10}{\GeV}$.
    Oscillation parameters are taken from NuFit~\cite{nufit} and given in Eq.~\ref{eq:PINGUparams}.}\label{fig:2_vs_3}
\end{figure}

In Fig.~\ref{fig:2_vs_3}, we have plotted the oscillation probabilities for an incoming $\nm$ with \si{\GeV} energy,
traveling 12000 km. The black lines are the probabilities of $\nm$ oscillation, after it has traversing through the diameter of the Earth.
The red lines are the probabilities of $\nm$ oscillation, but in the two-neutrino picture using Eq.~\ref{eq:2gen}. So in the left-most panel,
we are only considering the mixing between $\ne$ and $\nm$, completely neglecting $\theta_{13}$ and $\theta_{23}$.

The pattern is similar to when we compared $3\nu$ matter and vacuum oscillations: at $\SI{10}{\GeV}$, the two pictures converge.
It is evident that the full $3\nu$ picture is required for single-digit \si{\GeV} energies, and that mixings between all three flavors here are 
abundant. Increasing with energy, however, the influence of the $\nu_1$ mass eigenstate on the probabilities is reduced.
Thus, we are able to approximate resonant behavior as described in Eq.~\ref{eq:MSW} above $\SI{100}{\GeV}$ by only considering 
mixing between two neutrino species. We will utilize this observation later on to study a particular resonance in the \si{\TeV} range.
